---
title: "Survey Use May Influence Survey Response"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(summarytools)
library(sjmisc)
library(patchwork)
library(expss)
library(sandwich)
library(lmtest)
library(DescTools)
library(lubridate)
```

Note: The data attached is the raw data which includes the survey items in the second row.  These must be deleted before analyzing the data.  I manually deleted these in Excel.

```{r}
Data <- read_excel("Survey Use Survey Response Dataset April 20 2020.xlsx")
```

Descriptives

```{r}
#Gender
frq(Data$Q7.14)
#Sexual Orientation
frq(Data$Q7.15)
#Grades
frq(Data$Q7.16)
#Race/Ethnicity
frq(Data$Q7.17)
#Income
frq(Data$Q7.18)
```

Create an indicator for whether or not they responded honestly

```{r}
Data$P1_Honest<- Data %>% 
  mutate(P1_Honest=ifelse(Q3.4=="Answer honestly" & 
                            Q3.5=="Answer honestly" &
                            Q3.6=="Answer honestly" &
                            Q3.7=="Answer honestly" &
                            Q3.8=="Answer honestly",
                          "Honest","Dishonest")) %>% 
  pull()

Data$P2_Honest <- Data %>% 
  mutate(P2_Honest=ifelse(Q4.4=="Answer honestly" & 
                            Q4.5=="Answer honestly" &
                            Q4.6=="Answer honestly" &
                            Q4.7=="Answer honestly" &
                            Q4.8=="Answer honestly",
                          "Honest","Dishonest")) %>% 
  pull()

Data$P3_Honest <- Data %>% 
  mutate(P3_Honest=ifelse(Q5.4=="Answer honestly" & 
                            Q5.5=="Answer honestly" &
                            Q5.6=="Answer honestly" &
                            Q5.7=="Answer honestly" &
                            Q5.8=="Answer honestly",
                          "Honest","Dishonest")) %>% 
  pull()

Data$P4_Honest <- Data %>% 
  mutate(P4_Honest=ifelse(Q6.4=="Answer honestly" & 
                            Q6.5=="Answer honestly" &
                            Q6.6=="Answer honestly" &
                            Q6.7=="Answer honestly" &
                            Q6.8=="Answer honestly",
                          "Honest","Dishonest")) %>% 
  pull()
```

Time to transpose. Select the predictors (the responses to the DRI's) and the indicators we just created.  Also keep their participant ID.

```{r}
#Subset the data

df1 <- Data %>% 
  select(ResponseId,
         Q3.10:Q3.15,
         Q4.10:Q4.15,
         Q5.10:Q5.15,
         Q6.10:Q6.15,
         P1_Honest:P4_Honest)

df2 <- df1 %>% 
  unite(col="Profile_1",Q3.10:Q3.15,P1_Honest) %>% 
  unite(col="Profile_2",Q4.10:Q4.15,P2_Honest) %>%
  unite(col="Profile_3",Q5.10:Q5.15,P3_Honest) %>%
  unite(col="Profile_4",Q6.10:Q6.15,P4_Honest) %>% 
  pivot_longer(cols=Profile_1:Profile_4,names_to = "Response") %>% 
  separate(col="value",into=c("Comfort","Honest","Affect","Trust","Use","Other","Outcome"),sep="_") %>% 
  select(-ResponseId,-Response)

freq(df2$Outcome)
```

Run baseline linear model:

```{r}
#df2$Outcome[df2$Outcome=="NA"]<-NA

df2$Outcome <- factor(df2$Outcome)

class(df2$Outcome)

df2$Outcome2<-as.numeric(df2$Outcome,na.rm=TRUE)

freq(df2$Outcome2)

#Use SPSS package to assign value labels
df2$Outcome2[df2$Outcome2==2]<-0
val_lab(df2$Outcome2)=num_lab("1 Dishonest
                                     0 Honest")
val_lab(df2$Outcome2)

#Run baseline model
baseline<-glm(Outcome2~Honest,data=df2,family=binomial(link="logit"))
summary(baseline)
coeftest(baseline,vcov = vcovHC(baseline, type = "HC1"))

#save predicted values from baseline model
Predb<-baseline$fit

#append values to dataset
df2$Predb<-Predb

#categorize as pass/fail
df2$Predb1 <- NA
df2$Predb1[df2$Predb<.5]<-0
df2$Predb1[df2$Predb>=.5]<-1

#cross tabs (row,column)
mytable1<-table(df2$Outcome2,df2$Predb1)
mytable1
prop.table(mytable1,1)
dt<-as.data.frame.matrix(mytable1)
```

Prep variables for final model and graphing

```{r}
#Label NA's as NA
#df2$Comfort[df2$Comfort=="NA"] <- NA
#df2$Honest[df2$Honest=="NA"] <- NA
#df2$Affect[df2$Affect=="NA"] <- NA
#df2$Trust[df2$Trust=="NA"] <- NA
#df2$Use[df2$Use=="NA"] <- NA
#df2$Other[df2$Other=="NA"] <- NA

#Need to establish factor order for the x-axis of bar graph
df2$Comfort <- factor(df2$Comfort,
                      levels=c("Yes, to all", "Mostly", "Sort of", "No, not really"))
df2$Honest <- factor(df2$Honest,
                     levels=c("Yes","No"))

df2$Affect[df2$Affect=="Positively affect me"] <- "Positively"
df2$Affect[df2$Affect=="Negatively affect me"] <- "Negatively"
df2$Affect[df2$Affect=="Not affect me"] <- "No effect"
df2$Affect <- factor(df2$Affect,
                     levels=c("Positively","No effect","Negatively","Not sure"))
df2$Trust <- factor(df2$Trust,
                    levels=c("Yes","Mostly","Somewhat","No","Prefer not to say"))
df2$Use <- factor(df2$Use,
                  levels=c("Yes","Mostly","Somewhat","No","I don't know"))

df2$Other[df2$Other=="All of them"] <- "All"
df2$Other[df2$Other=="Most of them"] <- "Most"
df2$Other[df2$Other=="Some of them"] <- "Some"
df2$Other <- factor(df2$Other,
                    levels=c("All","Most","Some","Hardly any"))

#Look at the variables
frq(df2$Comfort)
frq(df2$Honest)
frq(df2$Affect)
frq(df2$Trust)
frq(df2$Use)
frq(df2$Other)
```

Run linear model for final model

```{r}
#Run model with all DRI's
binomial<-glm(Outcome2~Comfort+Honest+Affect+Trust+Use+Other,data=df2,family=binomial(link="logit"))
summary(binomial)
coeftest(binomial,vcov = vcovHC(binomial, type = "HC1"))

#save predicted values from binomial (full model)
Predf<-binomial$fitted.values

#append values to dataset
df2$Predf<-Predf

#categorize as pass/fail
df2$Predf1 <- NA
df2$Predf1[df2$Predf<.5]<-0
df2$Predf1[df2$Predf>=.5]<-1

#cross tabs (row,column)
mytable<-table(df2$Outcome2,df2$Predf1)
mytable
prop.table(mytable,1)
as.data.frame.matrix(mytable)

#Compute R2 for glms
PseudoR2(baseline, which = "Nagelkerke")
PseudoR2(binomial, which = "Nagelkerke")
```

Odds ratios:

```{r}
df2$Outcome <- factor(df2$Outcome,
                      levels = c("Honest","Dishonest"))

frq(df2$Outcome)

#Honest: 271
#Dishonest: 293

#Honest
df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Honest) %>%
  summarise(n=n())

#Comfort
df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Comfort) %>%
  summarise(n=n())

#Use
df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Use) %>% 
  summarise(n=n())

#Trust
df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Trust) %>% 
  summarise(n=n())

#Affect
df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Affect) %>% 
  summarise(n=n())

#Other
df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Other) %>% 
  summarise(n=n())
```


Create proportions for bar charts:

```{r}
df2$Outcome <- factor(df2$Outcome,
                      levels = c("Honest","Dishonest"))
pc<- df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Comfort) %>%
  summarise(n=n()) %>% 
  mutate(Proportion=n/sum(n)) %>% 
  ggplot(aes(x=Comfort,y=Proportion,fill=Outcome))+
  geom_bar(stat='identity',position='dodge')+
  ylim(0,1)+
  scale_fill_manual(values=c("#C4C4C4","#808080"),
                    breaks=c("Honest","Dishonest"),
                    labels=c("Participant responded honestly",
                             "Participant responded dishonestly"))+
  theme(panel.background = element_rect("white"),
        axis.line.x.bottom = element_line("black"),
        axis.line.y.left = element_line("black"),
        legend.title = element_blank(),
        legend.position = "none")+
  labs(y="")

ph <- df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Honest) %>% 
  summarise(n=n()) %>% 
  mutate(Proportion=n/sum(n)) %>% 
  ggplot(aes(x=Honest,y=Proportion,fill=Outcome))+
  geom_bar(stat='identity',position=position_dodge(preserve = "single"),width = .5)+
  ylim(0,1)+
  labs(legend="")+
  scale_fill_manual(values=c("#C4C4C4","#808080"),
                    breaks=c("Honest","Dishonest"),
                    labels=c("Participant responded honestly",
                             "Participant responded dishonestly"))+
  theme(panel.background = element_rect("white"),
        axis.line.x.bottom = element_line("black"),
        axis.line.y.left = element_line("black"),
        legend.title = element_blank(),
        legend.position = "none")+
  labs(y="")

pa<- df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Affect) %>%
  summarise(n=n()) %>% 
  mutate(Proportion=n/sum(n)) %>% 
  ggplot(aes(x=Affect,y=Proportion,fill=Outcome))+
  geom_bar(stat='identity',position='dodge')+
  ylim(0,1)+
  scale_fill_manual(values=c("#C4C4C4","#808080"),
                    breaks=c("Honest","Dishonest"),
                    labels=c("Participant responded honestly",
                             "Participant responded dishonestly"))+
  theme(panel.background = element_rect("white"),
        axis.line.x.bottom = element_line("black"),
        axis.line.y.left = element_line("black"),
        legend.title = element_blank(),
        legend.position = "none")+
  labs(y="")

pt<- df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Trust) %>%
  summarise(n=n()) %>% 
  mutate(Proportion=n/sum(n)) %>% 
  ggplot(aes(x=Trust,y=Proportion,fill=Outcome))+
  geom_bar(stat='identity',position='dodge')+
  ylim(0,1)+
  scale_fill_manual(values=c("#C4C4C4","#808080"),
                    breaks=c("Honest","Dishonest"),
                    labels=c("Participant responded honestly",
                             "Participant responded dishonestly"))+
  theme(panel.background = element_rect("white"),
        axis.line.x.bottom = element_line("black"),
        axis.line.y.left = element_line("black"),
        legend.title = element_blank(),
        legend.position = "none")+
  labs(y="")

pu<- df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Use) %>%
  summarise(n=n()) %>% 
  mutate(Proportion=n/sum(n)) %>% 
  ggplot(aes(x=Use,y=Proportion,fill=Outcome))+
  geom_bar(stat='identity',position='dodge')+
  ylim(0,1)+
  scale_fill_manual(values=c("#C4C4C4","#808080"),
                    breaks=c("Honest","Dishonest"),
                    labels=c("Participant responded honestly",
                             "Participant responded dishonestly"))+
  theme(panel.background = element_rect("white"),
        axis.line.x.bottom = element_line("black"),
        axis.line.y.left = element_line("black"),
        legend.title = element_blank(),
        legend.position = "bottom")+
  labs(y="")

po<- df2 %>% 
  na.omit() %>% 
  group_by(Outcome,Other) %>%
  summarise(n=n()) %>% 
  mutate(Proportion=n/sum(n)) %>% 
  ggplot(aes(x=Other,y=Proportion,fill=Outcome))+
  geom_bar(stat='identity',position='dodge')+
  ylim(0,1)+
  scale_fill_manual(values=c("#C4C4C4","#808080"),
                    breaks=c("Honest","Dishonest"),
                    labels=c("Participant responded honestly",
                             "Participant responded dishonestly"))+
  theme(panel.background = element_rect("white"),
        axis.line.x.bottom = element_line("black"),
        axis.line.y.left = element_line("black"),
        legend.title = element_blank(),
        legend.position = "bottom")+
  labs(y="")

ph + pc + pa + pt + pu + po + 
  plot_layout(ncol=2,guides="collect")&
  theme (legend.position="bottom")
```

Relationship between use and responses (figure)

```{r}
df3 <- Data %>% 
  select(Q7.5:Q7.7) %>% 
  gather(key="Variable",value="Response")

df3$Response<-factor(df3$Response,levels=c("No, never", "Rarely", "Occasionally", "Fairly often", "Yes, regularly"))

df3$Variable[df3$Variable=="Q7.5"] <- "When you take surveys, do you think about how the survey results will be used?"
df3$Variable[df3$Variable=="Q7.6"] <- "Do you ever choose not to take a survey, stop taking a survey, or skip survey items because you don't like the way your answers will be used?"
df3$Variable[df3$Variable=="Q7.7"] <- "Do you ever alter your survey answers to protect yourself because you don't like the way your answers will be used?"

mytab <- df3 %>%
  na.omit()%>%
  group_by(Variable)%>%
  count(Response) %>%
  mutate(prop = prop.table(n))
mytab

#reverse = true to re-order bars on graph
ggplot(mytab, aes(x = Variable, y = prop, fill = Response)) +
  geom_bar(position = position_stack(reverse=TRUE), stat = "identity", width = .4) +
  labs(y="Proportion", x="",col="")+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 50))+
  theme_classic()+
  scale_fill_grey(start = 0.85, end = 0.3)+ 
  theme(legend.title = element_blank())+
  theme(legend.position="bottom", legend.spacing.x = unit(.2, 'cm'))+
  theme(axis.text.y=element_text(size=12))+
  theme(axis.title.x=element_text(size=10, color="grey35"))+
  coord_flip()
```

Motivations for dishonest response

```{r}
df <- Data

df$Purpose <- df$Q7.12_1
df$Compensation <- df$Q7.12_2
df$Interest <- df$Q7.12_3
df$Use <- df$Q7.12_4
df$Anonymity <- df$Q7.12_5
df$HappenYou <- df$Q7.12_6
df$HappenFF <- df$Q7.12_7
df$HappenOther <- df$Q7.12_8
df$Info <- df$Q7.12_9
df$Benefit <- df$Q7.12_10
df$Trust <- df$Q7.12_11
df$Always <- df$Q7.12_12
df$Incriminate <- df$Q7.12_13
df$Other <- df$Q7.12_14

df$Purpose <- factor(df$Purpose)
df$Compensation <- factor(df$Compensation)
df$Interest <- factor(df$Interest)
df$Use <- factor(df$Use)
df$Anonymity <- factor(df$Anonymity)
df$HappenYou <- factor(df$HappenYou)
df$HappenFF <- factor(df$HappenFF)
df$HappenOther <- factor(df$HappenOther)
df$Info <- factor(df$Info)
df$Benefit <- factor(df$Benefit)
df$Trust <- factor(df$Trust)
df$Always <- factor(df$Always)
df$Incriminate <- factor(df$Incriminate)
df$Other <- factor(df$Other)

#subset the data
df_sub <- df %>% 
  select(Q7.11:Q7.11_4,Purpose:Other)

#Turn wide to long

df5 <- df_sub %>% 
  pivot_longer(cols = Q7.11:Q7.11_4, names_to = "Reason") %>% 
  select(-Reason)

df5$Reason <- df5$value

df5 <- df5 %>% select(-value)

df5$Reason <- factor(df5$Reason)

frq(df5$Reason)

df5 %>% 
  filter(!is.na(Reason)) %>% 
  group_by(Reason) %>% 
  summarise(Count=n()) %>% 
  mutate(Prop=Count/141) %>% 
  arrange(-Prop) 

df6 <- df_sub %>% 
  pivot_longer(cols = Purpose:Other, names_to = "Motivation") %>% 
  select(-Q7.11,-Q7.11_1,-Q7.11_2,-Q7.11_3,-Q7.11_4) %>% 
  pivot_wider(names_from = value, values_fn = list(value=length)) %>% 
  select(-"NA")

df7 <- df6 %>% 
  replace(is.na(.),0) %>% 
  mutate(Total=rowSums(.[2:6]),
         Prop=Total/141) %>% 
  arrange(-Total) %>% 
  select(Motivation,Prop,Never,Rarely,Sometimes,Usually,Always,Total)


##COUNT:

df4a <- df7

#To turn count into proprotion, multiply by .7092

100/141

df4a <- df4a %>% 
  mutate(Never=Never*.7092/100,
         Rarely=Rarely*.7092/100,
         Sometimes=Sometimes*.7092/100,
         Usually=Usually*.7092/100,
         Always=Always*.7092/100)

#Need to manually add in the 19 ppl who said yes to when anonymous bc 
#they didn't get asked the how often questions

df4a[14,8]=19*.7092/100
df4a[14,2]=19/141
df4a[14,3]=19*.7092/100

df5a <- df4a %>% 
  mutate(Prop=round(Prop,4)) %>% 
  mutate(Never=na_if(Never,"NaN"),
         Rarely=na_if(Rarely,"NaN"),
         Sometimes=na_if(Sometimes,"NaN"),
         Usually=na_if(Usually,"NaN"),
         Always=na_if(Always,"NaN"))

df6a <- df5a %>% 
  mutate(Motivation=replace(Motivation,Motivation=="Anonymity","Anonymity of my responses"),
         Motivation=replace(Motivation,Motivation=="HappenYou","Assurance that nothing bad will happen to you"),
         Motivation=replace(Motivation,Motivation=="Trust","Trust in the organization giving the survey"),
         Motivation=replace(Motivation,Motivation=="Use","Specific explanation of how survey results will be used"),
         Motivation=replace(Motivation,Motivation=="Incriminate","The survey questions don't incriminate me"),
         Motivation=replace(Motivation,Motivation=="Interest","Personal interest in the survey topic"),
         Motivation=replace(Motivation,Motivation=="HappenFF","Assurance that nothing bad will happen to your friends or family"),
         Motivation=replace(Motivation,Motivation=="Purpose","Specific explanation of survey purpose"),
         Motivation=replace(Motivation,Motivation=="Compensation","Financial compensation or course credit"),         
         Motivation=replace(Motivation,Motivation=="HappenOther","Assurance that the nothing bad will happen to people who share your race/ethnicity, sexual orientation, gender, or economic class"),                            
         Motivation=replace(Motivation,Motivation=="Benefit","Survey results are used to benefit survey takers"),                   
         Motivation=replace(Motivation,Motivation=="Info","Links to more information about the survey topic or researchers"),                   
         Motivation=replace(Motivation,Motivation=="Always","I always answer honestly no matter what"))

df7a <- df6a %>% 
  pivot_longer(cols=Never:Always,
               names_to = "Response",
               values_to = "Count")

df7a$Motivation <- factor(df7a$Motivation,
                         levels=c("Anonymity of my responses",
                                  "Assurance that nothing bad will happen to you",
                                  "Trust in the organization giving the survey",
                                  "The survey questions don't incriminate me",
                                  "Specific explanation of how survey results will be used",
                                  "Assurance that nothing bad will happen to your friends or family",
                                  "Personal interest in the survey topic",
                                  "Specific explanation of survey purpose",
                                  "Financial compensation or course credit",
                                  "Assurance that the nothing bad will happen to people who share your race/ethnicity, sexual orientation, gender, or economic class",
                                  "Survey results are used to benefit survey takers",
                                  "Links to more information about the survey topic or researchers",
                                  "Other",
                                  "I always answer honestly no matter what"))

df7a$Response <- factor(df7a$Response,
                       levels=c("Never","Rarely","Sometimes","Usually","Always"))

df7a %>% 
  ggplot(aes(x=fct_rev(Motivation),y=Count,fill=Response))+
  geom_bar(position = position_stack(reverse=TRUE), stat="identity")+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 66))+
  theme_classic()+
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        axis.text.y=element_text(size=10),
        axis.title.x=element_text(size=9, color="grey35"))+
  ylim(0,1)+
  scale_fill_grey(start=.85,end=.3)+
  labs(y="Proportion",x="")+
  #geom_text(aes(Motivation,Prop,label=gsub("0\\.", "\\.",Prop)), size=3,
           # hjust=.3)+
  coord_flip()
```

Did everyone lie?

```{r}
df3 <- df1 %>% 
  unite(col="Profile_1",Q3.10:Q3.15,P1_Honest) %>% 
  unite(col="Profile_2",Q4.10:Q4.15,P2_Honest) %>%
  unite(col="Profile_3",Q5.10:Q5.15,P3_Honest) %>%
  unite(col="Profile_4",Q6.10:Q6.15,P4_Honest) %>% 
  pivot_longer(cols=Profile_1:Profile_4,names_to = "Response") %>% 
  separate(col="value",into=c("Comfort","Honest","Affect","Trust","Use","Other","Outcome"),sep="_") %>% 
  select(-Response)

df4 <- df3 %>% 
  count(ResponseId,Outcome) %>% 
  group_by(ResponseId) %>% 
  mutate(prop=prop.table(n))

df5 <- df4 %>% 
  group_by(Outcome) %>% 
  count(prop)

(141-12)/141
```

